---
author: "Eleanor Fewings"
output: html_document
---

# start_section  
```{r start_section}

# Time stamp
Sys.time()

# Clean up
rm(list=ls())

# Load libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(Seurat)
library(cowplot)
library(stringr)
library(igraph)
library(ComplexHeatmap)
library(ggraph)
library(networkD3)

# Folders
setwd("~/Saez/Maurice_stent_analysis/")

```


## Load data
```{r load}

load("results/integration.de.RData")

```

#Clean up unnecessary data
```{r clean}

rm(list=c(ls(pattern="^stent"), ls(pattern="^avg")))

```


#Label cell types
```{r}

types <- c("0"="B-cells",
           "1"="T-cells",
           "2"="T-cells",
           "3"="Fibroblast",
           "4"="Macrophages",
           "5"="VSMC",
           "6"="RBC",
           "7"="Endothelial",
           "8"="Endothelial",
           "9"="Unknown",
           "10"="Neuron")

data <- SetIdent(data, value="cluster")
data <- RenameIdents(data, types)

data$celltype <- Idents(data)

#Create meta cols for each comparison
#Immune vs vsmc/fibro
data$comp1 <- NA
data$comp1[data$cluster %in% c("0", "1", "2", "4")] <- "immune"
data$comp1[data$cluster %in% c("3", "5")] <- "vsmcfibro"

#endothelial vs vsmc/fibro
data$comp2 <- NA
data$comp2[data$cluster %in% c("7", "8")] <- "endo"
data$comp2[data$cluster %in% c("3", "5")] <- "vsmcfibro"

#Subset data
case <- SetIdent(data, value="cc")
case <- subset(case, idents="case")

#Subset data
control <- SetIdent(data, value="cc")
control <- subset(control, idents="control")
```

#Function to convert rat to human genes - case
```{r}

rna.mx <- case@assays$RNA@data %>% as.matrix()

convertratGeneList <- function(x){
require("biomaRt")
human <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
rat <- useMart("ensembl", dataset = "rnorvegicus_gene_ensembl")
mouse <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")
genesV2 <-  getLDS(attributes = c("rgd_symbol"), filters = "rgd_symbol", values = x , mart = rat, attributesL = c("hgnc_symbol"), martL = human, uniqueRows=T)
humanx <- unique(genesV2[, 2])
return(genesV2)
}

#Create table of converted names
rat <- row.names(rna.mx)

converted <- convertratGeneList(rat)

#Reduce to names that have homolog
case.hg.mx <- rna.mx[row.names(rna.mx) %in% converted$RGD.symbol,]

#Filter gene names to match
converted.r <- converted[converted$RGD.symbol %in% row.names(rna.mx) & !duplicated(converted$RGD.symbol),]

converted.r <- converted.r[order(match(converted.r$RGD.symbol, row.names(case.hg.mx))),]

#Check if the two align
sum(row.names(case.hg.mx) != converted.r$RGD.symbol )

#Add human gene names
row.names(case.hg.mx) <- converted.r$HGNC.symbol


```

#Function to convert rat to human genes - control
```{r}

rna.mx <- control@assays$RNA@data %>% as.matrix()

#Create table of converted names
rat <- row.names(rna.mx)

converted <- convertratGeneList(rat)

#Reduce to names that have homolog
control.hg.mx <- rna.mx[row.names(rna.mx) %in% converted$RGD.symbol,]

#Filter gene names to match
converted.r <- converted[converted$RGD.symbol %in% row.names(rna.mx) & !duplicated(converted$RGD.symbol),]

converted.r <- converted.r[order(match(converted.r$RGD.symbol, row.names(control.hg.mx))),]

#Check if the two align
sum(row.names(control.hg.mx) != converted.r$RGD.symbol )

#Add human gene names
row.names(control.hg.mx) <- converted.r$HGNC.symbol


```

#Create format for cellphonedb
```{r}

meta.case <- case@meta.data[, colnames(case@meta.data) %in% c("cc", "comp1", "celltype")] %>% subset(select=c(celltype, comp1, cc))
meta.control <- control@meta.data[, colnames(control@meta.data) %in% c("cc", "comp2", "celltype")] %>% subset(select=c(celltype, comp2, cc))

#export
write.table(case.hg.mx, "scripts/cellphonedb/count.matrix.case.txt", quote=FALSE, sep="\t")
write.table(meta.case, "scripts/cellphonedb/metadata.case.txt", quote=FALSE, sep="\t")

write.table(control.hg.mx, "scripts/cellphonedb/count.matrix.control.txt", quote=FALSE, sep="\t")
write.table(meta.control, "scripts/cellphonedb/metadata.control.txt", quote=FALSE, sep="\t")

#Clean up data
rm(converted, converted.r, rna.mx, expr.plot, grid, npcs, p1, p2, rat, convertratGeneList)

```

#Load cellphoneDB output
```{r}

deconv.case <- read.table("scripts/cellphonedb/case_out/deconvoluted.txt", sep="\t", stringsAsFactors = FALSE, header=TRUE, check.names=FALSE)
means.case <- read.table("scripts/cellphonedb/case_out/means.txt", sep="\t", stringsAsFactors = FALSE, header=TRUE, check.names=FALSE)
pval.case <- read.table("scripts/cellphonedb/case_out/pvalues.txt", sep="\t", stringsAsFactors = FALSE, header=TRUE, check.names=FALSE)
smeans.case <- read.table("scripts/cellphonedb/case_out/significant_means.txt", sep="\t", stringsAsFactors = FALSE, header=TRUE, check.names=FALSE)

#Set rownames
rownames(means.case) <- means.case$id_cp_interaction
rownames(pval.case) <- pval.case$id_cp_interaction
rownames(smeans.case) <- smeans.case$id_cp_interaction

#Work with significant interactions
df <- smeans.case[,1:12]

smeans.case <- smeans.case[,13:ncol(smeans.case)]
pval.case <- pval.case[,12:ncol(pval.case)]
means.case <- means.case[,12:ncol(means.case)]


```

#Count number of interactions between populations
```{r}

grs <- unique(unlist(sapply(colnames(pval.case), function(z) strsplit(z, split="\\|")[[1]],simplify = FALSE)))

### Counts of interaction types between populations #####
cnt_interactions <- function(xx, grs) {
  CNT <- colSums(xx< 0.05)
  MAT <- matrix(0, ncol=length(grs), nrow=length(grs), dimnames=list(grs,grs))
  for(case in names(CNT)) {
    cnt <- CNT[case]
    gr <- strsplit(case, split="\\|")[[1]]
    MAT[gr[1], gr[2]] <- MAT[gr[1], gr[2]] + cnt
    MAT[gr[2], gr[1]] <- MAT[gr[2], gr[1]] + cnt
  }
  return(MAT)
}

MAT <- cnt_interactions(xx=pval.case, grs)


#--- Directional through ligands to receptor (removing receptor x receptor and ligand x ligand)
cnts_sub1 <- colSums(pval.case[df$receptor_a=="True" & df$receptor_b=="False",] < 0.05)
cnts_sub2 <- colSums(pval.case[df$receptor_a=="False" & df$receptor_b=="True",] < 0.05)

# We have to revert their names to be sum with the previous
names(cnts_sub2) <- sapply(names(cnts_sub2),function(z) paste(rev(strsplit(z,split="\\|")[[1]]),collapse = "|"))
cnts_sub2 <- cnts_sub2[names(cnts_sub1)] # same order
# Add both directionals
cnts <- cnts_sub1 + cnts_sub2
RECxLIG <- matrix(NA, ncol=length(grs), nrow=length(grs), dimnames=list(grs,grs))
for(case in names(cnts)) {
  cnt <- cnts[case]
  gr <- strsplit(case, split="\\|")[[1]]
  RECxLIG[gr[1], gr[2]] <- cnt
}
rm(cnt, cnts, cnts_sub1, cnts_sub2, gr, case)

```

## Landscape visualization
Color schema
```{r color_schemas}
colors5 <- c("#69a75b",
             "#b75fb3",
             "#bc8d3d",
             "#6885d0",
             "#cc5658")
names(colors5) <- as.character(unique(meta.case$celltype))
```

### Graph of connections
```{r CD10negative_cellphoneDB_graph, fig.width=11.5, fig.height=8.5, dpi=300}

#create an igraph network object from the weighted adjacency matrix stored in pc
net = igraph::graph_from_adjacency_matrix(MAT, weighted = TRUE)
#remove multiple edges (meaning keep only one connection between each two cell clusters)

net = igraph::simplify(net, edge.attr.comb = "max")

#Some information to use in our plots
Num.Interactions = E(net)$weight

Strength = strength(net)

#plot network with ggraph
set.seed(113)
lay = ggraph::create_layout(net, layout = "fr")
gg <- ggraph(lay) + 
  geom_edge_link(aes(color=Num.Interactions)) + 
  scale_edge_colour_gradient(  low = "#cacfcf",
                               high = "#b52002",
                               space = "Lab",
                               na.value = "grey50",
                               guide = "edge_colourbar") +
  geom_node_point(aes(size = Strength)) + 
  geom_node_text(aes(label = grs), repel=TRUE, size=6) +
  theme(panel.background = element_blank(),
        legend.text = element_text(size=14),
        legend.title = element_text(size=14))
print(gg)
```

## Heatmap of directional lig-rec interactions
```{r extended_data_Fig7a, fig.width=8.5, fig.height=4.5, dpi=300}

Heatmap(RECxLIG, name="Num",col=c("lightgrey","red"),
        row_title = "     Receptor", row_names_side = "left", row_title_gp = gpar(fontsize=22), show_row_dend = FALSE, show_column_dend = FALSE,
        column_title = "Source of ligand release", column_names_side = "top", column_title_gp = gpar(fontsize=22))

```

# Count sending and recieving interactions
```{r}
cnts_senders <- colSums(pval.case[df$receptor_a=="True" & df$receptor_b=="False",] < 0.05) %>% as.data.frame()
cnts_recievers <- colSums(pval.case[df$receptor_a=="False" & df$receptor_b=="True",] < 0.05) %>% as.data.frame()

#Set column names
colnames(cnts_senders) <- "count.send"
colnames(cnts_recievers) <- "count.rec"

#select senders and recievers from rownames
cnts_senders$sender <- sapply(strsplit(row.names(cnts_senders),"\\|"), `[`, 1)
cnts_recievers$recievers <- sapply(strsplit(row.names(cnts_recievers),"\\|"), `[`, 2)

cnts_senders <- cnts_senders %>% group_by(sender) %>% mutate(count=sum(count.send)) %>% select(sender, count) %>% unique()
cnts_recievers <- cnts_recievers %>% group_by(recievers) %>% mutate(count=sum(count.rec)) %>% select(recievers, count) %>% unique()

#Order from largest to smallest
cnts_senders <- cnts_senders[order(cnts_senders$count, decreasing = FALSE),]
cnts_senders$sender <- factor(cnts_senders$sender, levels=cnts_senders$sender)

cnts_recievers <- cnts_recievers[order(cnts_recievers$count, decreasing = FALSE),]
cnts_recievers$recievers <- factor(cnts_recievers$recievers, levels=cnts_recievers$recievers)

#Bar plots of senders and recievers
ggplot(cnts_senders, aes(x=sender, y=count)) +
  geom_bar(stat="identity") +
  coord_flip()

ggplot(cnts_recievers, aes(x=recievers, y=count)) +
  geom_bar(stat="identity") +
  coord_flip()


```


# Create Sankey plot
```{r}

#Create data for Sankey
#sigint <- 

p <- sankeyNetwork(Links = Energy$links, Nodes = Energy$nodes, Source = "source",
              Target = "target", Value = "value", NodeID = "name",
              units = "TWh", fontSize = 12, nodeWidth = 30)

```

